An elaborate data set on human gait and the effect of mechanical perturbations

### 2015-01-13 paper.tex
Added major section introductions, fixes #91.
- Added intros to the Methods and Results sections.
- Moved the markers subsection in as a paragraph in the mocap.txt section.
- Fixed figure captions to say red instead of purple.
- Got rid of last "velocity" mentions and changed to "speed".
- Moved wording to "commanded belt speed" instead of "input speed".

\section*{Methods}
\subsection*{Participants}
%
Fifteen able bodied subjects including four females and eleven males with an
average age of $24\pm4$ years, height of $1.75\pm0.09$ m, mass of $74\pm13$ kg
participated in the study. The study was approved by the Institutional Review
Board of Cleveland State University (\# 29904-VAN-HS) and written informed
consent was obtained from all participants. The data has been anonymized with
respect to the participants' identities and a unique identification number was
assigned to each subject. A selection of the meta data collected for each
subject is shown in Table~\ref{tab:subjects}.
%
\begin{table}
  \cprotect\caption{Information about the 15 study participants. The final
    three columns provide the trial numbers associated with each nominal
    treadmill speed. The measured mass is computed from the mean total vertical
    ground reaction force just after the calibration pose event, if possible.
    If the mass is reported without an accompanying standard deviation, it is
    the subject's self-reported mass. Additional trials found in the data set
    with a subject identification number 0 are trials with no subject, i.e.
    unloaded trials that can be used for inertial compensation purposes, and
    are not shown in the table. Generated by \verb|src/subject_table.py|.}
  \centering
  \small
  \input{tables/subjects.tex}
  \label{tab:subjects}
\end{table}

@@

The MATLAB script and Simulink model produce a comma-delimited text file of
five desired belt speed signals indexed by the time stamp. There are slow, normal,
and fast walking speeds and slow and fast running speeds.~\footnote{The running
signals were not used in the experiments presented in this paper.} The measured
speed of the treadmill belts from unloaded trials (79, 80, 81) are compared to
these commanded treadmill control input signals in
Figure~\ref{fig:input_output} to show the effect of the treadmill and
controller dynamics. The system introduces a delay and seems to act as a low
pass filter. The standard deviations of the measured speeds do not
significantly differ from the desired values:
\SIlist{0.05;0.12;0.2}{\meter\per\second} for the three speeds, respectively.
%
\begin{figure}
  \centering
  \includegraphics{figures/input_vs_output.pdf}
  \caption{Treadmill belt speed input signals (purple) and recorded output
    speeds (blue) for average belt speeds of
    \SIlist{0.8;1.2;1.6}{\meter\per\second}, respectively.}
  \label{fig:input_output}
\end{figure}

To show the effects of the treadmill dynamics and give an idea of the frequency
content of the actual perturbations, spectral density plots were created
by averaging a spectrogram of a twenty second Hamming window, shown
in Figure~\ref{fig:freq_analysis}. For all speeds, the frequency content of the
commanded and measured time series show similarity below 4~\si{\hertz} and
attenuation in the measured spectral density above 4~\si{\hertz}.
%
\begin{figure}
  \centering
  \includegraphics{figures/frequency_analysis.pdf}
  \caption{Frequency spectrum of the treadmill belt velocity input signal
    (purple) and the recorded output velocity (blue) for average belt speeds of
    0.8, 1.2, and 1.6 m/s, respectively}
  \label{fig:freq_analysis}
\end{figure}

@@

\section*{Results}
\subsection*{Raw Data}
%
The raw data consists of a set of ASCII tab delimited text files output from
both the ``mocap'' and ``record'' modules in D-Flow in addition to a manually
generated YAML file that contains all of the necessary meta data for the given
trial. These three files are stored in a hierarchy of directories with one
trial per directory. The directories are named in the following fashion
\verb+T001/+ where \verb+T+ stands for ``trial'' and the following three digits
are provide a unique trial identification number.

@@

Each trial directory contains a meta data file in the YAML format named in the
following style \verb|meta-xxx.yml| where \verb|xxx| is the three digit trial
identification number. There are three main headings in the file: \verb+study+,
\verb+subject+, and \verb+trial+. An example meta data file is shown in Listing
\ref{lis:example-meta-data}.

\subsection*{Markers}

@@

@@

For more insight into the difference in the unperturbed and perturbed data,
Figure~\ref{fig:gait-cycle-stats-comparison} compares the distribution of a few
gait cycle statistics. One can see that the perturbed strides have a much
larger variation in frequency and length.
%
\begin{figure}
  \centering
  \includegraphics{figures/unperturbed-perturbed-boxplot-comparison.pdf}
  \cprotect\caption{Box plots of the average belt speed, stride frequency, and
    stride length which compare unperturbed (purple) and perturbed (blue) gait
    cycles. The median is given with the box bounding the first and third
    quartiles and the whiskers bound the range of the data. Produced by
    \verb|src/unperturbed_perturbed_comparison.py|.}
  \label{fig:gait-cycle-stats-comparison}
\end{figure}

>>>>

\section*{Methods}
%
In this section, we describe our experimental design beginning with
descriptions of the participants and equipment. This is then followed by the
protocol details and specifics on the perturbation design.

\subsection*{Participants}
%
Fifteen able bodied subjects including four females and eleven males with an
average age of $24\pm4$ years, height of $1.75\pm0.09$ m, mass of $74\pm13$ kg
participated in the study. The study was approved by the Institutional Review
Board of Cleveland State University (\# 29904-VAN-HS) and written informed
consent was obtained from all participants. The data has been anonymized with
respect to the participants' identities and a unique identification number was
assigned to each subject. A selection of the meta data collected for each
subject is shown in Table~\ref{tab:subjects}.
%
\begin{table}
  \cprotect\caption{Information about the 15 study participants. The final
    three columns provide the trial numbers associated with each nominal
    treadmill speed. The measured mass is computed from the mean total vertical
    ground reaction force just after the calibration pose event, if possible.
    If the mass is reported without an accompanying standard deviation, it is
    the subject's self-reported mass. Additional trials found in the data set
    with a subject identification number 0 are trials with no subject, i.e.
    unloaded trials that can be used for inertial compensation purposes, and
    are not shown in the table. Generated by \verb|src/subject_table.py|.}
  \centering
  \small
  \input{tables/subjects.tex}
  \label{tab:subjects}
\end{table}

@@

The MATLAB script and Simulink model produce a comma-delimited text file of
five desired belt speed signals indexed by the time stamp. There are slow, normal,
and fast walking speeds and slow and fast running speeds.~\footnote{The running
signals were not used in the experiments presented in this paper.} The measured
speed of the treadmill belts from unloaded trials (79, 80, 81) are compared to
these commanded treadmill control input signals in
Figure~\ref{fig:input_output} to show the effect of the treadmill and
controller dynamics. The system introduces a delay and seems to act as a low
pass filter. The standard deviations of the measured speeds do not
significantly differ from the desired values:
\SIlist{0.05;0.12;0.2}{\meter\per\second} for the three speeds, respectively.
%
\begin{figure}
  \centering
  \includegraphics{figures/input_vs_output.pdf}
  \caption{Commanded treadmill belt speed (red) and the recorded
    speed (blue) for average belt speeds of
    \SIlist{0.8;1.2;1.6}{\meter\per\second}, respectively.}
  \label{fig:input_output}
\end{figure}

To show the effects of the treadmill dynamics and give an idea of the frequency
content of the actual perturbations, spectral density plots were created
by averaging a spectrogram of a twenty second Hamming window, shown
in Figure~\ref{fig:freq_analysis}. For all speeds, the frequency content of the
commanded and measured time series show similarity below 4~\si{\hertz} and
attenuation in the measured spectral density above 4~\si{\hertz}.
%
\begin{figure}
  \centering
  \includegraphics{figures/frequency_analysis.pdf}
  \caption{Power spectral density of the commanded treadmill belt speed
    (red) and the recorded speed (blue) for average belt speeds of
    \SIlist{0.8;1.2;1.6}{\meter\per\second}, respectively}
  \label{fig:freq_analysis}
\end{figure}

@@

\section*{Results}
%
Here we present some basic results. We first provide a detailed description of
the raw data followed by an overview of several computed variables that give an
idea of the characteristics of both the unperturbed and perturbed gait.

\subsection*{Raw Data}
%
The raw data consists of a set of ASCII tab delimited text files output from
both the ``mocap'' and ``record'' modules in D-Flow in addition to a manually
generated YAML file that contains all of the necessary meta data for the given
trial. These three files are stored in a hierarchy of directories with one
trial per directory. The directories are named in the following fashion
\verb+T001/+ where \verb+T+ stands for ``trial'' and the following three digits
are provide a unique trial identification number.

@@

@@

Each trial directory contains a meta data file in the YAML format named in the
following style \verb|meta-xxx.yml| where \verb|xxx| is the three digit trial
identification number. There are three main headings in the file: \verb+study+,
\verb+subject+, and \verb+trial+. An example meta data file is shown in
Listing~\ref{lis:example-meta-data}.

@@

For more insight into the difference in the unperturbed and perturbed data,
Figure~\ref{fig:gait-cycle-stats-comparison} compares the distribution of a few
gait cycle statistics. One can see that the perturbed strides have a much
larger variation in frequency and length.
%
\begin{figure}
  \centering
  \includegraphics{figures/unperturbed-perturbed-boxplot-comparison.pdf}
  \cprotect\caption{Box plots of the average belt speed, stride frequency, and
    stride length which compare unperturbed (red) and perturbed (blue) gait
    cycles. The median is given with the box bounding the first and third
    quartiles and the whiskers bound the range of the data. Produced by
    \verb|src/unperturbed_perturbed_comparison.py|.}
  \label{fig:gait-cycle-stats-comparison}
\end{figure}
